**EXPLORE & SUMMARIZE DATA | White Wine Quality Analysis**    
Aurore Dupont  
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Upload libraries needed
library(ggplot2)
library(reshape2)
library(gridExtra)
library(psych)
library(purrr)
library(tidyr)
library(memisc)
library(wesanderson)
library(scales)
library(knitr)
```

# Introduction  

In this project, we will be exploring a dataset about **the quality of white**
**wines**, using exploratory data analysis techniques to explore relationships 
in one variable to multiple variables in R.  

This dataset is public available for research. The details are described in 
[Cortez et al., 2009].  
P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis.  
Modeling wine preferences by data mining from physicochemical properties.  
In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.  

Available at:  
- [Elsevier](http://dx.doi.org/10.1016/j.dss.2009.05.016)  
- [Pre-press (pdf)](http://www3.dsi.uminho.pt/pcortez/winequality09.pdf)  
- [bib](http://www3.dsi.uminho.pt/pcortez/dss09.bib)  


### Descriptions of the variables

**Input variables** *(based on physicochemical tests)*:  
- **Fixed acidity** *(tartaric acid - g / dm^3)*: Most acids involved with wine 
or fixed or nonvolatile (do not evaporate readily)  
- **Volatile acidity** *(acetic acid - g / dm^3)*: The amount of acetic acid in 
wine, which at too high of levels can lead to an unpleasant, vinegar taste  
- **Citric acid** *(g / dm^3)*: Found in small quantities, citric acid can add 
‘freshness’ and flavor to wines  
- **Residual sugar** *(g / dm^3)*: The amount of sugar remaining after 
fermentation stops, it’s rare to find wines with less than 1 gram/liter and 
wines with greater than 45 grams/liter are considered sweet  
- **Chlorides** *(sodium chloride - g / dm^3)*: The amount of salt in the wine   
- **Free sulfur dioxide** *(mg / dm^3)*: The free form of SO2 exists in 
equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it 
prevents microbial growth and the oxidation of wine  
- **Total sulfur dioxide** *(mg / dm^3)*: Amount of free and bound forms of S02; 
in low concentrations, SO2 is mostly undetectable in wine, but at free SO2 
concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine   
- **Density** *(g / cm^3)*: The density of water is close to that of water 
depending on the percent alcohol and sugar content  
- **pH**: Describes how acidic or basic a wine is on a scale from 0 (very 
acidic) to 14 (very basic); most wines are between 3-4 on the pH scale  
- **Sulphates** *(potassium sulphate - g / dm3)*: A wine additive which can 
contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and
antioxidant  
- **Alcohol** *(% by volume)*: The percent alcohol content of the wine  

**Output variable** *(based on sensory data)*:  
- **Quality**: Score between 0 and 10  

Several of the attributes may be correlated, thus it makes sense to apply some 
sort of feature selection.  

------
  
# Univariate Plots Section  

We are starting with some preliminary exploration of the dataset.  
Summaries of the data and univariate plots will allow us to understand the 
structure of the individual variables in the dataset.  

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load the data
wine <- read.csv('wineQualityWhites.csv')

# Show what the dataset look like
head(wine, 3)
```

- Data have been loaded and the first lines have been displayed to know what the
dataset looks like.  

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Retrieve the dimension of the dataset
dim(wine)
```

- It includes 4,898 observations of 13 variables.  

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Return name of the variables
names(wine)
```

- Here are the names of each variables (descriptions can be found in the 
introduction).  

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Show the internal structure of the dataframe
str(wine)
```
- This part shows how the dataframe is structured: either integer or numerical 
data.  

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get basic descriptive statistics
describe(wine)
```

- We can notice that most of medians and means are close except for 
residual.sugar.  

```{r echo=FALSE, message=FALSE, warning=FALSE}
#colnames(wine)
#rownames(wine)
#levels(wine$ph)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Add a column name into the first column/first row
names(wine)[1] <- "wine_id"
head(wine, 3)
```

- We decided to rename the first column instead having X. New name: wine_id.  

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create a new column in which we can see the rank of the wines for 
#a better readibility:
# 1 - 3: Low
# 4 - 7: Medium
# 8 - 10: High
wine$level <- cut(wine$quality, c(1, 3, 7, 10), 
                                   labels = c("Low", "Medium", "High"))

head(wine, 25)
```

- We include a new column indicating if a wine has a Low, Medium, or High quality, 
such as:  
    - 1 - 3: Low  
    - 4 - 7: Medium  
    - 8 - 10: High  
  

**Now, let's plot some univariate data :**    

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}
theme_set(theme_minimal(5))

# Plot histograms based on log10 and continuous scales
p1 <- ggplot(aes(x = fixed.acidity), data = wine) +
  coord_cartesian(xlim = c(5.5, 8.5)) +
  geom_histogram() +
  scale_x_log10(breaks = seq(3.0, 15.0, 0.5)) +
  ggtitle("Tartaric Acid Distribution (log10)")

p2 <- ggplot(aes(x = volatile.acidity), data = wine) +
  coord_cartesian(xlim = c(0.1, 0.8)) +
  geom_histogram() +
  scale_x_log10(breaks = seq(0.0, 1.2, 0.1)) +
  ggtitle("Acetic Acid Distribution (log10)")

p3 <- ggplot(aes(x = citric.acid), data = wine) +
  coord_cartesian(xlim = c(0.10, 1.0)) +
  geom_histogram(binwidth = 0.05) +
  scale_x_log10(breaks = seq(0.0, 2.0, 0.2)) +
  ggtitle("Citric Acid Distribution (log10)")

p4 <- ggplot(aes(x = residual.sugar), data = wine) +
  coord_cartesian(xlim = c(0.6, 20)) +
  geom_histogram(binwidth = 0.05) +
  scale_x_log10(breaks = seq(0, 20, 2)) +
  ggtitle("Residual Sugar Distribution (log10)")

p5 <- ggplot(aes(x = chlorides), data = wine) +
  coord_cartesian(xlim = c(0.03, 0.07)) +
  geom_histogram(binwidth = 0.02) +
  scale_x_log10(breaks = seq(0.0, 0.35, 0.01)) +
  ggtitle("Sodium Chloride Distribution (log10)")

p6 <- ggplot(aes(x = free.sulfur.dioxide), data = wine) +
  coord_cartesian(xlim = c(5, 80)) +
  geom_histogram(binwidth = 0.05) +
  scale_x_log10(breaks = seq(2, 300, 15)) +
  ggtitle("Free Sulfur Dioxide Distribution (log10)")

p7 <- ggplot(aes(x = total.sulfur.dioxide), data = wine) +
  coord_cartesian(xlim = c(50, 300)) +
  geom_histogram(binwidth = 0.03) +
  scale_x_log10(breaks = seq(9, 440, 50)) +
  ggtitle("Total Sulfur Dioxide Distribution (log10)")

p8 <- ggplot(aes(x = density), data = wine) +
  coord_cartesian(xlim = c(0.9871, 1.003)) +
  geom_histogram(binwidth = 0.0002) +
  scale_x_log10(breaks = seq(0.9871, 1.0390, 0.02)) +
  ggtitle("Density Distribution (log10)")

p9 <- ggplot(aes(x = pH), data = wine) +
  coord_cartesian(xlim = c(2.8, 3,5)) +
  geom_histogram(binwidth = 0.005) +
  scale_x_log10(breaks = seq(2.5, 4.5, 0.3)) +
  ggtitle("pH Distribution (log10)")

p10 <- ggplot(aes(x = sulphates), data = wine) +
  coord_cartesian(xlim = c(0.3, 0.8)) +
  geom_histogram(binwidth = 0.02) +
  scale_x_log10(breaks = seq(0.0, 1.10, 0.2)) +
  ggtitle("Potassium Sulphate Distribution (log10)")

p11 <- ggplot(aes(x = alcohol), data = wine) +
  coord_cartesian(xlim = c(9, 13)) +
  geom_histogram(binwidth = 0.005) +
  scale_x_log10(breaks = seq(8, 14.5, 2)) +
  ggtitle("% of Alcohol by volume Distribution (log10)")

p12 <- ggplot(aes(x = quality), data = wine) +
  geom_bar() +
  ggtitle("Quality Distribution")

# Arrange plots in 3 columns
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol = 3)
```

- We plotted many histograms/barchart to see if some variables were standing out 
amongst all of them.  
- Log10 transformation have been performed to get a better visualization of the 
data.  
- At first glance, we can say that most of the variables have a fairly balanced
 distribution (despite some outliers) except for:  
        - Residual Sugar: It looks like that there are wines with a few sugar in 
        it and other with a lot of sugar.  
        - Alcohol: The distribution looks unusual.  
- This is confirmed by the summary data given above the histograms, when looking
at median and mean.  
- The normal distribution of each variables makes difficult to determine the ones 
that have a major impact to predict the quality.  
- Most of the wines are rated with a "Medium" quality.  
- Quality will be our main feature of interest.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=8}
# Univariate plot to check the outliers and statistics of the variables
subset(wine, select = -c(wine_id)) %>%
  keep(is.numeric) %>%
  gather() %>%
  ggplot(aes(x = '', y = value)) +
    facet_wrap(~ key, scales = "free", ncol = 3) +
    geom_boxplot() +
    theme(strip.background = element_rect(fill="aquamarine4"),
          strip.text = element_text(size = 10, face = "bold", color = "white"))
```

- The boxplots help to check the outliers of each variables. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Show summary statistics for variables  
summary(wine)
```

- As we have seen before, medians and means are quite close, except for residual 
sugar.  

# Univariate Analysis  

### What is the structure of your dataset?  

- The dataset consists of 4,898 observations of 14 variables (originally there 
were 13 variables but I created one more).  

### What is/are the main feature(s) of interest in your dataset?

- I am interested in looking at the quality of wine, *i.e. what chemicals* 
*characteristics are most important in predicting the quality*.  
- We can see that the wines are mostly rated from 3 to 9. Each expert graded the
wine quality between 0 (very bad) and 10 (very excellent). However, only a few 
are very bad or excellent.  

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

- Acidity like pH, alcohol, density and sugar are characteristics that might help 
to dig further into the analysis.  
- For example, we can ask whether wines with higher alcoholic content receive 
better ratings, whether sweeter wines contain a higher volume of alcohol or 
receive better ratings, but also what level of acidity (pH) is associated with 
the highest quality.  

### Did you create any new variables from existing variables in the dataset?

- A new variable called "level" has been created for a better 
readibility of the quality of wine. We can now easily identify if a wine has a 
bad or high quality.  

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?  

- I created histograms for each variables based on a log10 scale (except for the 
Quality) to help identify the distribution better.  

------

# Bivariate Plots Section

**As mentioned in the previous section, there might be some relationships between** 
**variables. To look at the correlation, we are creating a matrix.**   

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}
library(corrplot)

# Remove wine_id and level column to create a nice matrix
wine2 <- subset(wine, select = -c(wine_id, level))

# Attribute colors
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

# Create matrix
corrplot(cor(wine2), method = "color", col = col(200),
         type = "upper", order = "hclust", number.cex = .5,
         addCoef.col = "black", # Add coefficient of correlation
         tl.col = "black", tl.srt = 30, # Text label color and rotation
         # Combine with significance
         sig.level = 0.01, insig = "blank", 
         # Hide correlation coefficient on the principal diagonal
         diag = FALSE)
```

- The matrix allow us to identify right away what variables are the best to be 
analyzed.  
- We can notice a strong positive correlation between :  
        - residual sugar & density **(0.84)**,  
        - free sulfur dioxide & total sulfur dioxide **(0.62)**,  
        - density & total sulfur dioxide **(0.53)**,  
        - alcohol & quality **(0.44)**.  
- Most of the variables have a negative correlation with quality, which is our 
main feature of interest.  

```{r echo=FALSE, message=FALSE, warning=FALSE, Strength_of_Association}
# Guidelines to interpret correlation coefficient
corrcoef <- matrix(c(".1 to .3", "-0.1 to -0.3", ".3 to .5", "-0.3 to -0.5", 
                     ".5 to 1.0", "-0.5 to -1.0"), ncol = 2, byrow = TRUE)

colnames(corrcoef) <- c("Positive", "Negative")

rownames(corrcoef) <- c("Small", "Medium", "Large")

corrcoef <- as.table(corrcoef)

corrcoef
```
- This table help us interpret the correlation coefficient.  
- The strongness of the relationships highlighted above are confirmed.  

```{r echo=FALSE, message=FALSE, warning=FALSE}
#cor.test(wine$residual.sugar, wine$density, method = 'pearson')
#cor.test(wine$free.sulfur.dioxide, wine$total.sulfur.dioxide, method = 'pearson')
#cor.test(wine$density, wine$total.sulfur.dioxide, method = 'pearson')
#cor.test(wine$alcohol, wine$quality, method = 'pearson')
```


**Now, we are analizing the quality versus some variables.**  

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}
# Boxplot to help visualize the relationship between quality & other variables

# Quality vs. Alcohol
box1 <- ggplot(wine, aes(factor(quality), alcohol)) +
  geom_boxplot() +
  scale_fill_brewer(type = "seq") +
  stat_summary(fun.y = "mean", geom = "point", color = "orange", shape = 8, 
             size = 2) +
  ggtitle('Boxplot of % Alcohol by Quality')
  ggsave('QualityBoxPlot.png')
  
# Quality vs. Residual Sugar  
box2 <- ggplot(wine, aes(factor(quality), residual.sugar)) +
  geom_boxplot() +
  scale_fill_brewer(type = "seq") +
  scale_y_continuous(limits = c(0.0, 30.0)) +
  stat_summary(fun.y = "mean", geom = "point", color = "orange", shape = 8, 
             size = 2) +
  ggtitle('Boxplot of Residual Sugar by Quality')
  ggsave('QualityBoxPlot2.png')
  
# Quality vs. pH
box4 <- ggplot(wine, aes(factor(quality), pH)) +
  geom_boxplot() +
  scale_fill_brewer(type = "seq") +
  stat_summary(fun.y = "mean", geom = "point", color = "orange", shape = 8, 
             size = 2) +
  ggtitle('Boxplot of pH by Quality')
  ggsave('QualityBoxPlot3.png')

# Quality vs. Density    
box3 <- ggplot(wine, aes(factor(quality), density)) +
  geom_boxplot() +
  scale_fill_brewer(type = "seq") +
  scale_y_continuous(limits = c(0.985, 1.01)) +
  stat_summary(fun.y = "mean", geom = "point", color = "orange", shape = 8, 
             size = 2) +
  ggtitle('Boxplot of Density by Quality')
  ggsave('QualityBoxPlot4.png')
  
grid.arrange(box1, box2, box3, box4, ncol = 2)
```

- It seems that the wines with a high quality contain more alcohol.  
- However, the density shows an opposite trend. This proves a correlation between 
alcohol and density.  
- Although we can notice a good wine contains less sugar, its trend looks quite 
flat and doesn't indicate a major impact on the quality level.  
- The pH appears to be a bit greater for wines with a good quality.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}
# Residual Sugar vs. Density
plot1 <- ggplot(wine, aes(x = residual.sugar, y = density)) +
  geom_point(size = 1, shape = 23, alpha = 1/4, position = 'jitter') +
  xlim(0, 20) +
  ylim(0.99, 1.0) +
  ggtitle('Residual Sugar vs. Density')

# Free Sulfur Dioxide vs. Total Sulfur Dioxide
plot2 <- ggplot(wine, aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide)) +
  geom_point(size = 1, shape = 23, alpha = 1/4, position = 'jitter') +
  xlim(0, 100) +
  ylim(0, 300) +
  ggtitle('Free Sulfur Dioxide vs. Total Sulfur Dioxide')

# Density vs. Total Sulfur Dioxide
plot3 <- ggplot(wine, aes(x = density, y = total.sulfur.dioxide)) +
  geom_point(size = 1, shape = 23, alpha = 1/4, position = 'jitter') +
  xlim(0.99, 1) +
  ylim(0, 300) +
  ggtitle('Density vs. Total Sulfur Dioxide')

grid.arrange(plot1, plot2, plot3, ncol = 2)
```

- Here, we can see and confirm the relationship between:  
        - residual sugar & density,  
        - free sulfur dioxide & total sulfur dioxide,  
        - density & total sulfur dioxide,  
        - alcohol & quality.  
        
**Let's take a look at the relationships between the following variables:**  
        - alcohol & pH,  
        - alcohol & density,  
        - alcohol & residual sugar.  
        
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}
# Alcohol vs. pH
plot5 <- ggplot(aes(x = alcohol , y = pH), data = wine) +
  geom_point(size = 1, shape = 23, alpha = 1/3, position = 'jitter') +
  coord_cartesian(xlim = c(8, 15), ylim = c(2.70, 4.0)) +
  geom_smooth() +
  ggtitle('Alcohol vs. pH')

# Alcohol vs. Density
plot6 <-ggplot(wine, aes(x = alcohol , y = density)) +
  geom_point(size = 1, shape = 23, alpha = 1/3, position = 'jitter') +
  coord_cartesian(xlim = c(8, 15), ylim = c(0.985, 1.01)) +
  geom_smooth() +
  ggtitle('Alcohol vs. Density')

# Alcohol vs. Residual Sugar
plot7 <- ggplot(aes(x = alcohol , y = residual.sugar), data = wine) +
  geom_point(size = 1, shape = 23, alpha = 1/3, position = 'jitter') +
  coord_cartesian(xlim = c(8, 15), ylim = c(0.60, 17.0)) +
  geom_smooth() +
  ggtitle('Alcohol vs. Residual Sugar')

grid.arrange(plot5, plot6, plot7, ncol = 1)
```


- We can clearly see that the density decreases based on the volume of alcohol.  
- However, this trend is not noticable with the other variables, pH and residual 
sugar.  
- Alcohol vs. Ph scatterplot show that the regression line is quite steady. As 
for Alcohol vs. Residual sugar one, there a strong decrease in sugar for wines with 
8 to 10% of alcohol and then, the volume of sugar is steady as the volume of 
alcohol goes up.  

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?  

- Surprisingly, there is not many chemical characteristics that can predict the 
quality of wine, except for the percentage of alcohol.  
- Most of the variables have a negative correlation with quality.  
- It appears that a good wine contain a higher volume of alcohol and a higher pH,
but the concentration in residual sugar is low as well as the density.  

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

- We can notice a strong correlation between residual sugar & density, free 
sulfur dioxide & total sulfur dioxide, density & total sulfur dioxide, residual 
alcohol & quality.  

### What was the strongest relationship you found?

- The strong relationship between % of alcohol and quality cannot be denied.
- However, there is a strong relationship for residual sugar and density with a 
correlation coefficient of **0.84**.  

------

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
# Density vs. Alcohol across Level of quality
de.al <- ggplot(wine, aes(y = density, x = alcohol, 
                 color = as.factor(level))) +
  coord_cartesian(ylim = c(0.990, 1.0)) +
  geom_point(alpha = 1/3, size = 1) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_brewer(type='seq', palette='Dark2', 
                     guide = guide_legend(title = 'Quality'))

# Residual Sugar vs. Alcohol across Level
re.al <- ggplot(wine, aes(y = residual.sugar, x = alcohol, 
                 color = as.factor(level))) +
  coord_cartesian(ylim = c(0.60, 17.0)) +
  geom_point(alpha = 1/3, size = 1) +
  geom_smooth(method = "lm", se = FALSE, size = 1)  +
  scale_color_brewer(type = 'seq', palette='Set2', 
                     guide = guide_legend(title = 'Quality'))

grid.arrange(de.al, re.al, ncol = 1, 
             top = "Alcohol & Density vs. Alcohol & Residual Sugar Across Quality")
```

- Wines with a better quality and a higher volume of alcohol have less 
density and less sugar.  
- The gap between High and Medium wines is wider for thesugar content.  
- The density has a tendency to be lower accross all quality of wines as the 
volume of alcohol gets higher.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}
# Alcohol vs. pH across Quality
ggplot(wine, aes(x = alcohol, y = pH)) +
  geom_point(aes(color = factor(quality))) +
  scale_color_brewer(type = "seq") +
  facet_wrap(~quality, ncol = 7) +
  ggtitle('Alcohol vs. pH across Quality')
```

- The pH does not impact a lot the quality of the wine.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}
# Residual Sugar vs. Density vs. Level
ggplot(wine, aes(residual.sugar, density, color = level)) +
  geom_point(alpha = 2/3) +
  geom_smooth(method = "lm", se = FALSE) +
  coord_trans(x = "log10", 
              limx = c(min(wine$residual.sugar),
                       quantile(wine$residual.sugar, .999)),
              limy = c(min(wine$density),
                       quantile(wine$density, .999))) +
  scale_color_brewer(type = 'seq',
                     palette = 'YlOrBr',
                     guide = guide_legend(title = 'Quality',
                                          reverse = T,
                                          override.aes = list(alpha = 1, 
                                                              size = 2))) +
   ggtitle("Residual Sugar by Density and Quality")
```

- As seen before, wines with a low quality are sweeter and denser.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=2}
# pH vs. Density vs. Quality
ggplot(wine, aes(x=pH, y = density)) + 
  geom_smooth(method = 'lm') +
  coord_cartesian(ylim = c(0.985, 1.01)) +
  geom_point(alpha = 1/5, aes(color = quality)) + 
  scale_color_gradient(low = "yellow", high = "blue") +
  ggtitle("Density by pH and quality")
```

- Again, the pH doesn't show a major influence on the quality.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
# Density vs. Residual Sugar vs. Alcohol
ggplot(wine, aes(x = density, y = residual.sugar)) + 
  geom_point(aes(colour = alcohol), alpha = 0.35) +
  coord_cartesian(xlim = c(0.985, 1.005), ylim = c(0.60, 17.0)) +
  geom_smooth(method = "lm", aes(color = alcohol)) +
  ggtitle("Residual Sugar by Density & % of Alcohol") +
  ylab("Residual Sugar (g/L)") +
  xlab("Density") +
  scale_colour_continuous(name = "Volume of Alcohol (%)", low = "blue",
        high = "red")
```

- We can notice that low quality wines tend to be sweeter and denser, which 
corroborate our previous findings.  
- As a reminder, a "good" wine appears to have a higher volume of alcohol.  

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=5}
# Alcohol vs. Residual Sugar vs. pH
ggplot(wine, aes(x = alcohol, y = residual.sugar)) + 
  geom_point(aes(colour = pH), alpha = 0.35) +
  geom_smooth(method = "lm", aes(color = pH)) +
  ggtitle("Residual Sugar by Alcohol & pH") +
  ylab("Residual Sugar (g/L)") +
  xlab("Alcohol Concentration (%)") +
  ylim(0,25) +
  scale_colour_continuous(name = "pH", low = "blue",
        high = "red")
```

- We do not identify a strong relationship between pH vs. Alcohol & Sugar.  


**Now, let's take a look at our linear model.**  

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Linear Model Implementation
m1 <- lm(quality ~ alcohol, data = wine)
m2 <- update(m1, ~ . + sulphates)
m3 <- update(m2, ~ . + pH)
m4 <- update(m3, ~ . + density)
m5 <- update(m4, ~ . + free.sulfur.dioxide)
m6 <- update(m5, ~ . + total.sulfur.dioxide)
m7 <- update(m6, ~ . + chlorides)
m8 <- update(m7, ~ . + residual.sugar)
m9 <- update(m8, ~ . + citric.acid)
m10 <- update(m9, ~ . + volatile.acidity)
m11 <- update(m10, ~ . + fixed.acidity)


mtable(m1, m2, m3, m4, m5, m6, m7, m8, m9, m10, m11, sdigits = 3)
```

- The R-squared value is 0.282. This means that the variables would have an impact
of 28% only on the quality.  

```{r echo=FALSE, message=FALSE, warning=FALSE}
# thisWine = data.frame(quality = 7, density = 0.9974, residual.sugar = 3.4, 
                      # pH = 3.37, alcohol = 12, sulphates = 0.50)
# modelEstimate = predict(m11, newdata = thisWine, interval = "prediction", 
                        # level = .95)
# exp(modelEstimate)
```

# Multivariate Analysis

- Alcohol and Density have an impact on the quality of wine.  
- The density has a robust relationship with residual sugar.  
- We have been able to emphasize that quality increases as alcohol increases and
density & residual sugar decrease.  

------

# Final Plots and Summary

### Plot One | Quality vs. Alcohol  


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}
ggplot(wine, aes(x = level, y = alcohol,  fill = level)) + 
  geom_boxplot() + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", color = 'gold', 
               size = 1) +
  ylim(quantile(wine$alcohol, prob = 0.00), 
       quantile(wine$alcohol, prob = 0.99)) +
  ggtitle("Range of Alcohol per Quality Level") + 
  xlab("Quality") +
  ylab("Alcohol (%)") +
  theme(title = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 8), 
        axis.text = element_text(size = 8),
        legend.position = "none", panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = wes_palette(n = 3, name = "GrandBudapest1"))
```

### Description One

- Most of the wines are categorized with a medium quality.  
- This shows that wines with higher quality tend to have a higher volume of 
alcohol.  
- Alcohol is the variable that has the strongest correlation based on quality.  

### Plot Two | Quality vs. Alcohol & Quality vs. Density  


```{r echo=FALSE, message=FALSE, warning=FALSE}
finalbox1 <- ggplot(wine, aes(x = factor(quality), y = alcohol, fill = level)) +
  geom_jitter(alpha = .3, color = 'pink')  +
  geom_boxplot(alpha = .5, color = 'black')+
  stat_summary(fun.y = "mean", geom = "point", color = "green", size = 1) +
  labs(x = "Quality",
       y = "Alcohol (% by Volume)",
       title = "Alcohol vs. Quality") +
  theme(title = element_text(size = 12, face = 'bold'), 
                           panel.grid.major = element_blank(), 
                           panel.grid.minor = element_blank()) +
  scale_fill_manual(values = wes_palette(n = 3, name = "GrandBudapest1"))

finalbox2 <- ggplot(wine, aes(x = factor(quality), y = density, fill = level)) +
  geom_jitter(alpha = .3, color = 'pink')  +
  geom_boxplot(alpha = .5, color = 'black')+
  stat_summary(fun.y = "mean", geom = "point", color = "green", size = 1) +
  scale_y_continuous(limits = c(0.985, 1.01)) +
  labs(x = "Quality",
       y = "Density (g/mL)",
       title = "Density vs. Quality") +
  theme(title = element_text(size = 12, face = 'bold'), 
                           panel.grid.major = element_blank(), 
                           panel.grid.minor = element_blank()) +
  scale_fill_manual(values = wes_palette(n = 3, name = "GrandBudapest1"))


grid.arrange(finalbox1, finalbox2, ncol = 1)
```

### Description Two

- I want to show the negative correlation between alcohol and density and their
impact on the quality.  
- Based on the outliers and intervals, density and alcohol do not give better 
wine if we had to choose one variable only to predict the wine quality.  

### Plot Three | Residual Sugar vs. Density vs. Quality  


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=6}
ggplot(wine, aes(x = density, y = residual.sugar, color = as.factor(quality))) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Quality', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +
  theme_dark() +
  scale_x_continuous(limits = c(0.99, 1.004)) + 
  scale_y_continuous(trans = log10_trans()) +
  ggtitle('Residual Sugar (log10) by Density and Quality') +
  xlab("Density (g/mL)") +
  ylab("Residual Sugar (g/L)") +
  theme(title = element_text(size = 10, face = "bold"),
      axis.title = element_text(size = 8), 
      axis.text = element_text(size = 8),
      legend.text = element_text(size = 8), panel.grid.major = element_blank(), 
      panel.grid.minor = element_blank())
```

### Description Three

- This plot summarizes the fact that good wines contain less sugar and are less 
dense.  

------

# Reflection

- We went over a dataset containing 4,898 observations and 13 variables.  
- It is surprising that only a few factors (alcohol, density and residual sugar) 
affect the wine quality.  
- It is likely that with more variables - such as grape types, states, weather 
(rain and sun impact the quality), expertise, organic or not, fermentation time,
type of barrel,- a thorough analysis could have been conducted.  

- Where did I run into difficulties in the analysis?  
    - At first, it was not clear what variables I had to investigate. That's why
    I chose to plot all of them and see if one of them was standing out. Then, 
    the relationships between variables got clearer thanks to the correlation 
    matrix.
    - Also, I am not really familiar with chemical compounds, that is why I chose
    to dig further the following variables Density, pH, Residual Sugar and Alcohol. 
    These terms are more familiar to me.  

- Where did I find successes?  
    - I knew that the percentage of alcohol has an impact on the wine quality. 
    And indeed, it got confirmed through the analysis.  

# References
https://s3.amazonaws.com/content.udacity-data.com/courses/ud651/diamondsExample_2016-05.html  
http://rprogramming.net/rename-columns-in-r/  
http://www.cookbook-r.com/Manipulating_data/Adding_and_removing_columns_from_a_data_frame/  
https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set  
https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html  
https://stackoverflow.com/questions/21140798/error-using-corrplot/21141799  
http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r  
https://statistics.laerd.com/statistical-guides/pearson-correlation-coefficient-statistical-guide.php  
https://www.cyclismo.org/tutorial/R/tables.html  
http://r-statistics.co/Linear-Regression.html  
https://stackoverflow.com/questions/43359050/error-continuous-value-supplied-to-discrete-scale-in-default-data-set-example/43359104  
http://felixfan.github.io/ggplot2-remove-grid-background-margin/  
https://rstudio-pubs-static.s3.amazonaws.com/228019_f0c39e05758a4a51b435b19dbd321c23.html#1_plot_one_variable_-_x:_continuous_or_discrete

